GVN-Hoist: Hoisting Computations from Branches

### 2016-08-25 hoist.tex
add description of current implementation of code hoisting in llvm

  We present a code hoisting pass that we implemented in llvm. It is based on
  Global Value Numbering infrastructure available in llvm. The experimental
  results show an average of 2.5\% savings in code size, although the code size
  increases in many cases because it enables more inlining. This is an
  optimistic algorithm in the sense that we consider all identical computations
  in a function as potential candidates to be hoisted. We make an extra effort
  to hoist candidates by partitioning the potential candidates in a way to
  enable partial hoisting in case common hoisting points for all the candidates
  cannot be found. We also formalize cases when register pressure will reduce as
  a result of hoisting.
\end{abstract}

@@

Rosen et al. explain moving computations successors to remove redundancies
\cite{rosen1988global}. Their algorithm iterates on computations of same rank
and move the code with identical computations from the sibling branch. This is a
pessimistic approach (TODO: is it?). Also there is no notion of partially
hoisting the computations so their approach may result in missing many hoisting
opportunities.

A lot of bug reports in GCC and LLVM bugzillas
\cite{GCCCodeHoistingBugs,LLVMCodeHoistingBugs}, no papers on this topic yet.

>>>>

  We present a code hoisting pass that we implemented in LLVM. It is based on
  Global Value Numbering infrastructure available in LLVM. The experimental
  results show an average of 2.5\% savings in code size, although the code size
  increases in many cases because it enables more inlining. This is an
  optimistic algorithm in the sense that we consider all identical computations
  in a function as potential candidates to be hoisted. We make an extra effort
  to hoist candidates by partitioning the potential candidates in a way to
  enable partial hoisting in case common hoisting points for all the candidates
  cannot be found. We also formalize cases when register pressure will reduce as
  a result of hoisting.
\end{abstract}

@@

There are a lot of bug reports in GCC and LLVM bugzillas
\cite{GCCCodeHoistingBugs,LLVMCodeHoistingBugs}, showing the interest in having
a more powerful code hoist transform.  The current LLVM implementation of code
hoisting in SimplifyCFG.cpp $HoistThenElseCodeToIf()$ is very limited to
hoisting from identical basic blocks: the instructions of two sibling basic
blocks are read in the same time, and all the instructions of the blocks are
hoisted to the common parent block as long as the compiler is able to prove that
the instructions are equivalent.  This implementation does not allow for an easy
extension: first in terms of compilation time overhead the implementation would
become quadratic in number of instructions to bisimulate, and second the
equivalence of instructions currently uses $$I1->isIdenticalToWhenDefined(I2)$$
would need to be rewritten to be more general, leading to using a mechanism
similar to GVN.

Rosen et al. explain moving computations successors to remove redundancies
\cite{rosen1988global}. Their algorithm iterates on computations of same rank
and move the code with identical computations from the sibling branch. Also
there is no notion of partially hoisting the computations so their approach may
result in missing many hoisting opportunities.
